---
title: "Update Genus Master Taxa List"
author: "Emma Jones"
date: "1/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# built in R 3.6.2
library(tidyverse)
library(lubridate)
library(config)
library(pool)
library(pins)
library(readxl)
library(dbplyr)

# Server connection things
conn <- config::get("connectionSettings") # get configuration settings

board_register_rsconnect(key = conn$CONNECT_API_KEY,  #Sys.getenv("CONNECT_API_KEY"),
                         server = conn$CONNECT_SERVER)#Sys.getenv("CONNECT_SERVER"))

## For testing: connect to ODS production
pool <- dbPool(
  drv = odbc::odbc(),
  Driver = "ODBC Driver 11 for SQL Server",#Driver = "SQL Server Native Client 11.0",
  Server= "DEQ-SQLODS-PROD,50000",
  dbname = "ODS",
  trusted_connection = "yes"
)

source('helperFunctions/VSCI_metrics_GENUS.R')
source('helperFunctions/VCPMI_metrics_GENUS.R')

```

## Background

This document archives the steps taken to convert the SCI calculations from the original Genus level master taxa list to an updated Genus level master taxa list. This step is critical to the continued improvement of data collection to ensure biologists have the appropriate taxonomic resolution available when entering data, as well as improves the process for identifying the target taxonomic level for a given taxa. Completing this step will allow for a smoother transition (on the database side) to a genus level SCI after it is approved. 

Prior to this document, Drew Garey and Billy Van Wart manually QAed the existing Genus level master taxa list. They updated taxonomic names to the most current naming protocols, expanded the taxonomic list to species level where applicable (and indicated a target ID level), and updated habit/functional feeding group information to species level. These changes were reviewed by the statewide biologists workgroup and then passed along to Emma Jones to update the SCI calculation methods with these improvements.

This document will implement the proposed changes to the master taxa list and verify the updates do not change the existing Family level SCI scores across all SCI methods. 

### Genus Level Master Taxa List Archive

Before moving forward, we will preserve a snapshot of the existing Genus Level master taxa list from ODS via the R server pin.

```{r Genus master taxa list archive}
masterTaxaListGenus <- pin_get('ejones/masterTaxaGenus', board = 'rsconnect')
#saveRDS(masterTaxaListGenus, 'data/archive/masterTaxaGenus01032022.RDS')
```

### New Genus Level Master Taxa List

This is the final version from Drew and Billy.

```{r new genus master taxa list}
masterTaxaListGenusNew <- read_excel('data/masterTaxaGenus_01112022.xlsx', sheet = 'masterTaxaGenus') %>% # Emma update

  #read_excel('data/masterTaxaGenus_01062022.xlsx', sheet = 'masterTaxaGenus') %>% # Emma update
  #read_excel('data/masterTaxaGenus_27Sept2021.xlsx', sheet = 'masterTaxaGenus') %>%  # original from Drew/Billy
  mutate_at(c("FamTolVal", "TolVal", "GVSCI_TolVal"), as.numeric)
# replace character NA with NA
masterTaxaListGenusNew[masterTaxaListGenusNew == "NA"] <- NA
glimpse(masterTaxaListGenusNew)
glimpse(masterTaxaListGenus)

# what are the rows with changed FinalID?
filter(masterTaxaListGenusNew, ! FinalID %in% masterTaxaListGenus$FinalID)

# any changes between FinalID and OldFinalID columns?
mutate(masterTaxaListGenusNew, diff = ifelse(FinalID != OldFinalID, 1, 0)) %>% 
  arrange(desc(diff))


# any changes between FinalID and OldFinalID columns?
mutate(masterTaxaListGenusNew, diff = ifelse(FinalID != GVSCI_FinalID, 1, 0)) %>% 
  arrange(desc(diff))
```

So quick differences are 10 new rows, 29 changed FinalID's, two name changes between FinalID and OldFinalID, and 73 changes from FinalID to GVSCI_FinalID.


### Test VSCI calculation with new master taxa out of the box

Pull test benthic data and run VSCI to see if any changes

```{r test vsci prep}
# last check
masterTaxaGenus <- pool %>% tbl(in_schema("wqm",  "Edas_Benthic_Master_Taxa_View")) %>%
  as_tibble() %>%
  # make columns match expected format
  rename('Phylum' = 'PHYLUM_NAME',
         'Class' = 'CLASS_NAME',
         'Subclass' = 'SUBCLASS_NAME',
         'Order' = 'ORDER_NAME',
         'Suborder' = 'SUBORDER_NAME',
         'Superfamily' = 'SUPERFAMILY_NAME',
         'Family' = 'FAMILY_NAME',
         'Subfamily' = 'SUBFAMILY_NAME',
         'Tribe' = 'TRIBE_NAME',
          'Genus' = 'GENUS_NAME',
          'Species' = 'SPECIES_NAME',
          "Final VA Family ID" =  "WBMT_FINAL_FAMILY_ID",
          "FinalID" = "WBMT_FINAL_ID",
          "TolVal" = "WBMT_TOLERANCE_VALUE",
          "FFG" =   "FEEDING_GROUP", 
          "Habit" = "HABIT", 
          "FamFFG" =  "FAMILY_FEEDING_GROUP",
          "FamTolVal" = "WBMT_FAM_TOLERANCE_VALUE",
          "FamHabit" ="FAMILY_HABIT") %>% 
  dplyr::select(Phylum, Class, Subclass, Order, Suborder, Superfamily, Family, Subfamily, Tribe, 
                Genus, Species, `Final VA Family ID`, FinalID, TolVal, FFG, 
                Habit, FamFFG, FamTolVal, FamHabit) # keep EDAS Master Taxa list names

# just get fields needed for SCI
masterTaxaGenus <- masterTaxaListGenusNew %>% 
  dplyr::select(Phylum, Class, Subclass, Order, Suborder, Superfamily, Family, Subfamily, Tribe, 
                Genus, Species, `Final VA Family ID`, FinalID, TolVal, FFG, 
                Habit, FamFFG, FamTolVal, FamHabit) # keep EDAS Master Taxa list names

vmast <- masterTaxaGenus %>%
        # get Family level tolerance value, FFG
        rename('GenusTolVal' = 'TolVal',
               'TolVal' = 'FamTolVal',
               'GenusFFG' = 'FFG',
               'FFG' = 'FamFFG',
               'GenusHabit' = 'Habit',
               'Habit' = 'FamHabit') %>%
        mutate(e=ifelse(Order=="Ephemeroptera", 1, 0),
               p=ifelse(Order=="Plecoptera",1,0),
               t=ifelse(Order=="Trichoptera", 1, 0),
               tmin=ifelse((Order=="Trichoptera" & Family != "Hydropsychidae") | 
                             (Order=="Trichoptera" & is.na(Family)) , 1, 0), 
               ept=ifelse(e+p+t>0,1,0), 
               scraper = ifelse(FFG=="Scraper", 1, 0),
               chiro = ifelse(Family=="Chironomidae",1, 0),
               ptmin = ifelse(p + tmin > 0,1,0),
               `clinger-HS` = ifelse(Habit == 'Clinger' & ! Family %in% c("Hydropsychidae","Simuliidae"), 1, 0)) %>%
        # Then put it in long format so it can be merged to and input taxa list
        select(`Final VA Family ID`,TolVal, e,p,t, ept,ptmin, scraper, chiro,`clinger-HS`) %>% 
        distinct(`Final VA Family ID`, .keep_all = T) %>% # drop multiple rows bc working back to family level data from genus
        filter(!is.na(`Final VA Family ID`)) %>%
        pivot_longer(-`Final VA Family ID`, names_to = 'metric', values_to = 'metric_val') %>%
        #  pivot_longer(-`Final VA Family ID`, names_to = 'metric', values_to = 'metric_val') %>%
        filter(!is.na(metric_val))

### SCI calculation

SCI <- function(stationBenthicsDateRange, SCIchoice, benSamps, masterTaxaGenus, vmast){
  edas_options <- select(masterTaxaGenus, Class, Subclass, Order, Suborder, Superfamily, Family, `Final VA Family ID`, FinalID) %>%
    mutate(across(where(is.factor), as.character))
  edas_list <- select(edas_options, `Final VA Family ID`,FinalID)
  # for Excluding taxa, need a list of all Family level designations that may end up as a FinalID
  # these are all unique Family names and the few taxa that are the only 
  GenusNames <- c(unique(edas_options$FinalID)[!is.na(unique(edas_options$FinalID))])
  FamilyNames <- unique(edas_options$Family)[!is.na(unique(edas_options$Family))]
  SuperfamilyNames <- unique(edas_options$Superfamily)[!is.na(unique(edas_options$Superfamily))]
  SuborderNames <- unique(edas_options$Suborder)[!is.na(unique(edas_options$Suborder))]
  OrderNames <- unique(edas_options$Order)[!is.na(unique(edas_options$Order))]
  SubclassNames <- unique(edas_options$Subclass)[!is.na(unique(edas_options$Subclass))]
  ClassNames <- unique(edas_options$Class)[!is.na(unique(edas_options$Class))]
  
  
  EDASrare <- stationBenthicsDateRange %>%
    ########## #filter(str_detect(BenSampID, 'R110') & RepNum == 1) %>% # keep only rarified data and Rep1's
    mutate(Count = Individuals) %>% # Rename to match formatting of functions
    ######`Excluded Taxa` = ifelse(`Excluded Taxa` == T, -1, 0)) %>% 
    select(BenSampID, FinalID, Count, `Excluded Taxa`) %>%
    mutate(GenusTaxaLevel = ifelse(FinalID %in% GenusNames, T, F),
           FamilyTaxaLevel = ifelse(FinalID %in% FamilyNames, T, F),
           SuperfamilyTaxaLevel = ifelse(FinalID %in% SuperfamilyNames, T, F),
           SuborderTaxaLevel = ifelse(FinalID %in% SuborderNames, T, F),
           OrderTaxaLevel = ifelse(FinalID %in% OrderNames, T, F),
           SubclassTaxaLevel = ifelse(FinalID %in% SubclassNames, T, F),
           ClassTaxaLevel = ifelse(FinalID %in% ClassNames, T, F))
  
  # Work FinalID back up to Family Level
  EDASrare2 <- left_join(EDASrare,edas_list, by="FinalID") %>%
    filter(!is.na(`Final VA Family ID`)) %>%
    rename( `Genus Level Excluded Taxa` = `Excluded Taxa`)
  
  # We also need to do a little data manipulation to incorporate biologist exclusion information appropriately.
  exclusionMath  <- EDASrare2 %>%
    mutate(`Family Level Excluded Taxa` = 
             ifelse(`Genus Level Excluded Taxa` == -1, 
                    ifelse(`SuperfamilyTaxaLevel` == TRUE | `SuborderTaxaLevel` == TRUE | `OrderTaxaLevel` == TRUE | 
                             `SubclassTaxaLevel` == TRUE | `ClassTaxaLevel` == TRUE , -1, 0), 0 )) %>%
    # had to get super ifelse nesty here to make this logic work, ugly but works
    group_by(BenSampID, `Final VA Family ID`) %>%
    summarise(`Family Level Count` = sum(Count), 
              #`Genus Level Excluded Taxa` = sum(`Genus Level Excluded Taxa`),
              `Family Level Taxa` = n(),
              `Family Level Excluded Taxa` = sum(`Family Level Excluded Taxa`),
              `Final Family Level Taxa` = `Family Level Taxa` + sum(`Family Level Excluded Taxa`) )
  
  # Join bug traits
  bugTraits <- left_join(exclusionMath,vmast,by=c('Final VA Family ID') )
  
  
  if(SCIchoice == 'VSCI'){SCI <- VSCIcalculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VSCI',
           `SCI Threshold` = 60) %>% 
    rename("SCI Score" ="Fam SCI")}
  if(SCIchoice == 'VCPMI63 + Chowan'){SCI <- VCPMI63calculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VCPMI63 + Chowan',
           `SCI Threshold` = 40) %>% 
    rename("SCI Score" ="CPMI63+CHOWAN")}
  if(SCIchoice == 'VCPMI65 - Chowan'){SCI <- VCPMI65calculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VCPMI65 - Chowan',
           `SCI Threshold` = 40) %>% 
    rename("SCI Score" ="CPMI65-CHOWAN")}
  
  SCI <- left_join(SCI, benSamps, by = 'BenSampID')
  
  return(SCI) 
}
```


```{r benthics}
# pull benthics
benthics <- pool %>% tbl(in_schema("wqm",  "Edas_Benthic_View")) %>%
        as_tibble() %>%
        rename( "StationID" = "STA_ID",
                "BenSampID"  = "WBS_SAMP_ID",
                "RepNum" = "WBS_REP_NUM",
                "FinalID" = "WBMT_FINAL_ID",
                "Individuals" = "WBE_INDIVIDUALS",
                "ID Comments" = "WBE_COMMENT",
                "Entered By" = "WBE_INSERTED_BY", # not in EDAS table but good info
                "Taxonomist" = "TAXONOMIST_NAME",  # not in EDAS table but good info
                "Entered Date" = "WBE_INSERTED_DATE") %>%
        mutate(`Excluded Taxa` = ifelse(WBE_EXCLUDED_TAXA_YN == "Y", -1, 0)) %>%
        dplyr::select(StationID, BenSampID, RepNum, FinalID, Individuals, `Excluded Taxa`, `ID Comments`, Taxonomist, `Entered By`, `Entered Date`)
# for testing after changes to master taxa
# benthics1 <- benthics
# benthics <- benthics %>% 
#   mutate(FinalID = case_when(FinalID == 'Nixe' ~ "Afghanurus",
#                              FinalID == 'Leucrocuta/Nixe' ~ 'Leucrocuta/Afghanurus',
#                              TRUE ~ as.character(FinalID)))

#pull bensamps
benSamps <- pool %>% tbl(in_schema("wqm",  "Edas_Benthic_Sample_View")) %>%
        as_tibble() %>%
        # fix names
        rename( "StationID" = "STA_ID",
                "BenSampID"  = "WBS_SAMP_ID",
                "RepNum" = "WBS_REP_NUM",
                "Sample Comments" = "WBS_COMMENT",
                "Entered By" = "WBS_INSERTED_BY", # not in EDAS table but good info
                "Collected By" = "COLLECTOR_NAME",  # not in EDAS table but good info
                "Entered Date" = "WBS_INSERTED_DATE",
                "Gradient" = "WBCM_DESCRIPTION",
                "Taxonomist" = "TAXONOMIST_NAME",  # not in EDAS table but good info
                "Target Count" = "WBS_TARGET_COUNT",
                "Field Team" = "WBS_FIELD_TEAM",
                "Collection Date" = "FDT_DATE_TIME") %>%
        # Add sample season 
        mutate(monthday = as.numeric(paste0(sprintf("%02d",month(`Collection Date`)),
                                            sprintf("%02d",day(`Collection Date`)))),
               Season = case_when(monthday >= 0215 & monthday <= 0615 ~ 'Spring',
                                  monthday >= 0815 & monthday <= 1215 ~ 'Fall',
                                  TRUE ~ as.character("Outside Sample Window"))) %>%
        dplyr::select(StationID, BenSampID, RepNum, `Collection Date`, `Sample Comments`, `Collected By`, `Field Team`, `Entered By`,
                      Taxonomist, `Entered Date`, Gradient, `Target Count`, Season)
```

Update FinalID from GVSCI_final ID back to FinalID for analysis.

```{r gvsci_finalID to FinalID}
# #archive
# benthics1 <- benthics
# 
# benthicsNew <- rename(benthics, FinalID_CEDS= FinalID) %>% 
#  # mutate(n = row_number()) %>% 
#   #left_join(rename(benthics, FinalID_CEDS= FinalID), 
#   left_join(
#                        dplyr::select(masterTaxaListGenusNew, FinalID, GVSCI_FinalID),
#                       # by = c('FinalID_CEDS' = 'GVSCI_FinalID')) %>% 
#                        by = c('FinalID_CEDS' = 'FinalID')) %>% 
#   mutate(FinalID_ForSCI = FinalID_CEDS) %>% 
#   #mutate(FinalID_ForSCI = coalesce(FinalID, FinalID_CEDS)) %>% 
#   dplyr::select(StationID, BenSampID, RepNum, FinalID = FinalID_ForSCI,
#                 Individuals, `Excluded Taxa`, `ID Comments`, Taxonomist, 
#                 `Entered By`, `Entered Date`) 
# VSCIresults <- SCI(benthicsNew, "VSCI", benSamps,  masterTaxaGenus, vmast) %>%
#         mutate_if(is.numeric, round, digits=2) %>%# rounds all numeric columns to 2 decimal places
#         arrange(`Collection Date`)
# VSCIresults$Season <-  factor(VSCIresults$Season,levels=c("Spring","Outside Sample Window","Fall"))#,ordered=T)

```



```{r vsci calculation}
# run everyone through VSCI
VSCIresults <- SCI(benthics, "VSCI", benSamps,  masterTaxaGenus, vmast) %>%
        mutate_if(is.numeric, round, digits=2) %>%# rounds all numeric columns to 2 decimal places
        arrange(`Collection Date`)
VSCIresults$Season <-  factor(VSCIresults$Season,levels=c("Spring","Outside Sample Window","Fall"))#,ordered=T)

```
Compare to existing VSCI results.

```{r VSCI results comparison}
VSCIresultsPin <- pin_get("ejones/VSCIresults", board = "rsconnect")

comparison <- left_join(VSCIresultsPin, VSCIresults,  by = c("BenSampID",   "SCI", "SCI Threshold", "StationID", "RepNum", "Collection Date", "Sample Comments", "Collected By", "Field Team", "Entered By", "Taxonomist", "Entered Date", "Gradient", "Target Count", "Season")) %>% 
  mutate(difference = abs(`SCI Score.x` - `SCI Score.y`)) %>% 
  dplyr::select(sort(current_vars())) %>%  #arrange names alphabetically
  dplyr::select(difference, BenSampID, StationID, `Collection Date`, everything()) %>% 
  arrange(desc(difference))

nrow(filter(comparison, difference !=0)) # 41 issues, not bad
```

```{r}
# efficiently work through problems
knownIssues <- filter(benthics, FinalID %in% c('Nixe', 'Plectrocnemia')) %>% 
  distinct(BenSampID) %>% pull()
View(
filter(comparison, difference !=0) %>% 
  filter(BenSampID %in% knownIssues))
```


```{r}
benSamps1 <- benSamps

sampleID <- "LEV19577R110"
#'LIV8493R110' #Nixe
#'XMJ19555R110' #Plectrocnemia
#'GOO8492'  #Nixe
#'GOO8492R110' #Nixe

stationBenthicsDateRange<- filter(benthicsNew, BenSampID %in% sampleID) 
#stationBenthicsDateRange<- filter(benthics, BenSampID %in% sampleID) 
SCIchoice <- "VSCI"
benSamps <- filter(benSamps1, BenSampID %in% sampleID) 
#masterTaxaGenus, vmast, benSamps, masterTaxaGenus, vmast
```

```{r}

  
  EDASrare <- stationBenthicsDateRange %>%
    ########## #filter(str_detect(BenSampID, 'R110') & RepNum == 1) %>% # keep only rarified data and Rep1's
    mutate(Count = Individuals) %>% # Rename to match formatting of functions
    ######`Excluded Taxa` = ifelse(`Excluded Taxa` == T, -1, 0)) %>% 
    select(BenSampID, FinalID, Count, `Excluded Taxa`) %>%
    mutate(GenusTaxaLevel = ifelse(FinalID %in% GenusNames, T, F),
           FamilyTaxaLevel = ifelse(FinalID %in% FamilyNames, T, F),
           SuperfamilyTaxaLevel = ifelse(FinalID %in% SuperfamilyNames, T, F),
           SuborderTaxaLevel = ifelse(FinalID %in% SuborderNames, T, F),
           OrderTaxaLevel = ifelse(FinalID %in% OrderNames, T, F),
           SubclassTaxaLevel = ifelse(FinalID %in% SubclassNames, T, F),
           ClassTaxaLevel = ifelse(FinalID %in% ClassNames, T, F))
  
  # Work FinalID back up to Family Level
  EDASrare2 <- left_join(EDASrare,edas_list, by="FinalID") %>%
    filter(!is.na(`Final VA Family ID`)) %>%
    rename( `Genus Level Excluded Taxa` = `Excluded Taxa`)
  
  # We also need to do a little data manipulation to incorporate biologist exclusion information appropriately.
  exclusionMath  <- EDASrare2 %>%
    mutate(`Family Level Excluded Taxa` = 
             ifelse(`Genus Level Excluded Taxa` == -1, 
                    ifelse(`SuperfamilyTaxaLevel` == TRUE | `SuborderTaxaLevel` == TRUE | `OrderTaxaLevel` == TRUE | 
                             `SubclassTaxaLevel` == TRUE | `ClassTaxaLevel` == TRUE , -1, 0), 0 )) %>%
    # had to get super ifelse nesty here to make this logic work, ugly but works
    group_by(BenSampID, `Final VA Family ID`) %>%
    summarise(`Family Level Count` = sum(Count), 
              #`Genus Level Excluded Taxa` = sum(`Genus Level Excluded Taxa`),
              `Family Level Taxa` = n(),
              `Family Level Excluded Taxa` = sum(`Family Level Excluded Taxa`),
              `Final Family Level Taxa` = `Family Level Taxa` + sum(`Family Level Excluded Taxa`) )
  
  # Join bug traits
  bugTraits <- left_join(exclusionMath,vmast,by=c('Final VA Family ID') )
  
  
  if(SCIchoice == 'VSCI'){SCI <- VSCIcalculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VSCI',
           `SCI Threshold` = 60) %>% 
    rename("SCI Score" ="Fam SCI")}
  if(SCIchoice == 'VCPMI63 + Chowan'){SCI <- VCPMI63calculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VCPMI63 + Chowan',
           `SCI Threshold` = 40) %>% 
    rename("SCI Score" ="CPMI63+CHOWAN")}
  if(SCIchoice == 'VCPMI65 - Chowan'){SCI <- VCPMI65calculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VCPMI65 - Chowan',
           `SCI Threshold` = 40) %>% 
    rename("SCI Score" ="CPMI65-CHOWAN")}
  
  SCI <- left_join(SCI, benSamps, by = 'BenSampID')
```


Now test with VCPMI 63

Compare to existing results.

```{r VCPMI results comparison}
benSamps <- benSamps1 #reset

VCPMI63results <- SCI(benthics, "VCPMI63 + Chowan", benSamps,  masterTaxaGenus, vmast) %>%
        mutate_if(is.numeric, round, digits=2) %>%# rounds all numeric columns to 2 decimal places
        arrange(`Collection Date`)
VCPMI63results$Season <-  factor(VCPMI63results$Season,levels=c("Spring","Outside Sample Window","Fall"))#,ordered=T)

VCPMI63resultsPin <- pin_get("ejones/VCPMI63results", board = "rsconnect")

comparison <- left_join(VCPMI63resultsPin, VCPMI63results,  by = c("BenSampID",   "SCI", "SCI Threshold", "StationID", "RepNum", "Collection Date", "Sample Comments", "Collected By", "Field Team", "Entered By", "Taxonomist", "Entered Date", "Gradient", "Target Count", "Season")) %>% 
  mutate(difference = abs(`SCI Score.x` - `SCI Score.y`)) %>% 
  dplyr::select(sort(current_vars())) %>%  #arrange names alphabetically
  dplyr::select(difference, BenSampID, StationID, `Collection Date`, everything()) %>% 
  arrange(desc(difference))

nrow(filter(comparison, difference !=0)) # 35 issues, not bad
```
35 issues


```{r}
# efficiently work through problems
knownIssues <- filter(benthics, FinalID %in% c('Nixe', 'Plectrocnemia')) %>% 
  distinct(BenSampID) %>% pull()
View(
filter(comparison, difference !=0) %>% 
  filter(! BenSampID %in% knownIssues))
```


```{r}
benSamps1 <- benSamps

sampleID <- 'MTR19644R110'
#'LIV8493R110' #Nixe
#'XMJ19555R110' #Plectrocnemia
#'GOO8492'  #Nixe
#'GOO8492R110' #Nixe

stationBenthicsDateRange<- filter(benthics, BenSampID %in% sampleID) 
SCIchoice <- "VSCI"
benSamps <- filter(benSamps1, BenSampID %in% sampleID) 
#masterTaxaGenus, vmast, benSamps, masterTaxaGenus, vmast
```


```{r}

  
  EDASrare <- stationBenthicsDateRange %>%
    ########## #filter(str_detect(BenSampID, 'R110') & RepNum == 1) %>% # keep only rarified data and Rep1's
    mutate(Count = Individuals) %>% # Rename to match formatting of functions
    ######`Excluded Taxa` = ifelse(`Excluded Taxa` == T, -1, 0)) %>% 
    select(BenSampID, FinalID, Count, `Excluded Taxa`) %>%
    mutate(GenusTaxaLevel = ifelse(FinalID %in% GenusNames, T, F),
           FamilyTaxaLevel = ifelse(FinalID %in% FamilyNames, T, F),
           SuperfamilyTaxaLevel = ifelse(FinalID %in% SuperfamilyNames, T, F),
           SuborderTaxaLevel = ifelse(FinalID %in% SuborderNames, T, F),
           OrderTaxaLevel = ifelse(FinalID %in% OrderNames, T, F),
           SubclassTaxaLevel = ifelse(FinalID %in% SubclassNames, T, F),
           ClassTaxaLevel = ifelse(FinalID %in% ClassNames, T, F))
  
  # Work FinalID back up to Family Level
  EDASrare2 <- left_join(EDASrare,edas_list, by="FinalID") %>%
    filter(!is.na(`Final VA Family ID`)) %>%
    rename( `Genus Level Excluded Taxa` = `Excluded Taxa`)
  
  # We also need to do a little data manipulation to incorporate biologist exclusion information appropriately.
  exclusionMath  <- EDASrare2 %>%
    mutate(`Family Level Excluded Taxa` = 
             ifelse(`Genus Level Excluded Taxa` == -1, 
                    ifelse(`SuperfamilyTaxaLevel` == TRUE | `SuborderTaxaLevel` == TRUE | `OrderTaxaLevel` == TRUE | 
                             `SubclassTaxaLevel` == TRUE | `ClassTaxaLevel` == TRUE , -1, 0), 0 )) %>%
    # had to get super ifelse nesty here to make this logic work, ugly but works
    group_by(BenSampID, `Final VA Family ID`) %>%
    summarise(`Family Level Count` = sum(Count), 
              #`Genus Level Excluded Taxa` = sum(`Genus Level Excluded Taxa`),
              `Family Level Taxa` = n(),
              `Family Level Excluded Taxa` = sum(`Family Level Excluded Taxa`),
              `Final Family Level Taxa` = `Family Level Taxa` + sum(`Family Level Excluded Taxa`) )
  
  # Join bug traits
  bugTraits <- left_join(exclusionMath,vmast,by=c('Final VA Family ID') )
  
  
  if(SCIchoice == 'VSCI'){SCI <- VSCIcalculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VSCI',
           `SCI Threshold` = 60) %>% 
    rename("SCI Score" ="Fam SCI")}
  if(SCIchoice == 'VCPMI63 + Chowan'){SCI <- VCPMI63calculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VCPMI63 + Chowan',
           `SCI Threshold` = 40) %>% 
    rename("SCI Score" ="CPMI63+CHOWAN")}
  if(SCIchoice == 'VCPMI65 - Chowan'){SCI <- VCPMI65calculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VCPMI65 - Chowan',
           `SCI Threshold` = 40) %>% 
    rename("SCI Score" ="CPMI65-CHOWAN")}
  
  SCI <- left_join(SCI, benSamps, by = 'BenSampID')
```



















Now test with VCPMI

Compare to existing VSCI results.

```{r VCPMI results comparison}
benSamps <- benSamps1 #reset

VCPMI65results <- SCI(benthics, "VCPMI65 - Chowan", benSamps,  masterTaxaGenus, vmast) %>%
        mutate_if(is.numeric, round, digits=2) %>%# rounds all numeric columns to 2 decimal places
        arrange(`Collection Date`)
VCPMI65results$Season <-  factor(VCPMI63results$Season,levels=c("Spring","Outside Sample Window","Fall"))#,ordered=T)

VCPMI65resultsPin <- pin_get("ejones/VCPMI65results", board = "rsconnect")

comparison <- left_join(VCPMI65resultsPin, VCPMI65results,  by = c("BenSampID",   "SCI", "SCI Threshold", "StationID", "RepNum", "Collection Date", "Sample Comments", "Collected By", "Field Team", "Entered By", "Taxonomist", "Entered Date", "Gradient", "Target Count", "Season")) %>% 
  mutate(difference = abs(`SCI Score.x` - `SCI Score.y`)) %>% 
  dplyr::select(sort(current_vars())) %>%  #arrange names alphabetically
  dplyr::select(difference, BenSampID, StationID, `Collection Date`, everything()) %>% 
  arrange(desc(difference))

nrow(filter(comparison, difference !=0)) # 40 issues, not bad
```
40

```{r}
# efficiently work through problems
knownIssues <- filter(benthics, FinalID %in% c('Nixe', 'Plectrocnemia')) %>% 
  distinct(BenSampID) %>% pull()
View(
filter(comparison, difference !=0) %>% 
  filter(! BenSampID %in% knownIssues))
```


```{r}
benSamps1 <- benSamps

sampleID <- 
#'LIV8493R110' #Nixe
#'XMJ19555R110' #Plectrocnemia
#'GOO8492'  #Nixe
#'GOO8492R110' #Nixe

stationBenthicsDateRange<- filter(benthics, BenSampID %in% sampleID) 
SCIchoice <- "VSCI"
benSamps <- filter(benSamps1, BenSampID %in% sampleID) 
#masterTaxaGenus, vmast, benSamps, masterTaxaGenus, vmast
```

```{r}

  
  EDASrare <- stationBenthicsDateRange %>%
    ########## #filter(str_detect(BenSampID, 'R110') & RepNum == 1) %>% # keep only rarified data and Rep1's
    mutate(Count = Individuals) %>% # Rename to match formatting of functions
    ######`Excluded Taxa` = ifelse(`Excluded Taxa` == T, -1, 0)) %>% 
    select(BenSampID, FinalID, Count, `Excluded Taxa`) %>%
    mutate(GenusTaxaLevel = ifelse(FinalID %in% GenusNames, T, F),
           FamilyTaxaLevel = ifelse(FinalID %in% FamilyNames, T, F),
           SuperfamilyTaxaLevel = ifelse(FinalID %in% SuperfamilyNames, T, F),
           SuborderTaxaLevel = ifelse(FinalID %in% SuborderNames, T, F),
           OrderTaxaLevel = ifelse(FinalID %in% OrderNames, T, F),
           SubclassTaxaLevel = ifelse(FinalID %in% SubclassNames, T, F),
           ClassTaxaLevel = ifelse(FinalID %in% ClassNames, T, F))
  
  # Work FinalID back up to Family Level
  EDASrare2 <- left_join(EDASrare,edas_list, by="FinalID") %>%
    filter(!is.na(`Final VA Family ID`)) %>%
    rename( `Genus Level Excluded Taxa` = `Excluded Taxa`)
  
  # We also need to do a little data manipulation to incorporate biologist exclusion information appropriately.
  exclusionMath  <- EDASrare2 %>%
    mutate(`Family Level Excluded Taxa` = 
             ifelse(`Genus Level Excluded Taxa` == -1, 
                    ifelse(`SuperfamilyTaxaLevel` == TRUE | `SuborderTaxaLevel` == TRUE | `OrderTaxaLevel` == TRUE | 
                             `SubclassTaxaLevel` == TRUE | `ClassTaxaLevel` == TRUE , -1, 0), 0 )) %>%
    # had to get super ifelse nesty here to make this logic work, ugly but works
    group_by(BenSampID, `Final VA Family ID`) %>%
    summarise(`Family Level Count` = sum(Count), 
              #`Genus Level Excluded Taxa` = sum(`Genus Level Excluded Taxa`),
              `Family Level Taxa` = n(),
              `Family Level Excluded Taxa` = sum(`Family Level Excluded Taxa`),
              `Final Family Level Taxa` = `Family Level Taxa` + sum(`Family Level Excluded Taxa`) )
  
  # Join bug traits
  bugTraits <- left_join(exclusionMath,vmast,by=c('Final VA Family ID') )
  
  
  if(SCIchoice == 'VSCI'){SCI <- VSCIcalculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VSCI',
           `SCI Threshold` = 60) %>% 
    rename("SCI Score" ="Fam SCI")}
  if(SCIchoice == 'VCPMI63 + Chowan'){SCI <- VCPMI63calculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VCPMI63 + Chowan',
           `SCI Threshold` = 40) %>% 
    rename("SCI Score" ="CPMI63+CHOWAN")}
  if(SCIchoice == 'VCPMI65 - Chowan'){SCI <- VCPMI65calculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VCPMI65 - Chowan',
           `SCI Threshold` = 40) %>% 
    rename("SCI Score" ="CPMI65-CHOWAN")}
  
  SCI <- left_join(SCI, benSamps, by = 'BenSampID')
```


Last QA check, any taxa in the benthics table that are not in the new master taxa list? The metrics caught 2 but that doesnt mean it is totally inclusive of all options. These could have been purposefully changed or missed by accident.

```{r all missing taxa}
unique(benthics$FinalID)[! unique(benthics$FinalID) %in% masterTaxaListGenusNew$FinalID]
```

