---
title: "VSCI Calculation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#these are necessary to calculate from local data sources
library(tidyverse)
library(lubridate)
# these are necessary to pull data from ODS
library(DBI)
library(pool)
```


## Important functions

These run the VSCI/VCPMI

```{r}
source('helperFunctions/VSCI_metrics_GENUS.R')
source('helperFunctions/VCPMI_metrics_GENUS.R')

#and master SCI function that allows you to run whatever SCI you want easily
# SCI calculation

SCI <- function(stationBenthicsDateRange, SCIchoice, benSamps, masterTaxaGenus, vmast){
  edas_options <- select(masterTaxaGenus, Class, Subclass, Order, Suborder, Superfamily, Family, `Final VA Family ID`, FinalID) %>%
    mutate(across(where(is.factor), as.character))
  edas_list <- select(edas_options, `Final VA Family ID`,FinalID)
  # for Excluding taxa, need a list of all Family level designations that may end up as a FinalID
  # these are all unique Family names and the few taxa that are the only 
  GenusNames <- c(unique(edas_options$FinalID)[!is.na(unique(edas_options$FinalID))])
  FamilyNames <- unique(edas_options$Family)[!is.na(unique(edas_options$Family))]
  SuperfamilyNames <- unique(edas_options$Superfamily)[!is.na(unique(edas_options$Superfamily))]
  SuborderNames <- unique(edas_options$Suborder)[!is.na(unique(edas_options$Suborder))]
  OrderNames <- unique(edas_options$Order)[!is.na(unique(edas_options$Order))]
  SubclassNames <- unique(edas_options$Subclass)[!is.na(unique(edas_options$Subclass))]
  ClassNames <- unique(edas_options$Class)[!is.na(unique(edas_options$Class))]
  
  
  EDASrare <- stationBenthicsDateRange %>%
    ########## #filter(str_detect(BenSampID, 'R110') & RepNum == 1) %>% # keep only rarified data and Rep1's
    mutate(Count = Individuals) %>% # Rename to match formatting of functions
    ######`Excluded Taxa` = ifelse(`Excluded Taxa` == T, -1, 0)) %>% 
    select(BenSampID, FinalID, Count, `Excluded Taxa`) %>%
    mutate(GenusTaxaLevel = ifelse(FinalID %in% GenusNames, T, F),
           FamilyTaxaLevel = ifelse(FinalID %in% FamilyNames, T, F),
           SuperfamilyTaxaLevel = ifelse(FinalID %in% SuperfamilyNames, T, F),
           SuborderTaxaLevel = ifelse(FinalID %in% SuborderNames, T, F),
           OrderTaxaLevel = ifelse(FinalID %in% OrderNames, T, F),
           SubclassTaxaLevel = ifelse(FinalID %in% SubclassNames, T, F),
           ClassTaxaLevel = ifelse(FinalID %in% ClassNames, T, F))
  
  # Work FinalID back up to Family Level
  EDASrare2 <- left_join(EDASrare,edas_list, by="FinalID") %>%
    filter(!is.na(`Final VA Family ID`)) %>%
    rename( `Genus Level Excluded Taxa` = `Excluded Taxa`)
  
  # We also need to do a little data manipulation to incorporate biologist exclusion information appropriately.
  exclusionMath  <- EDASrare2 %>%
    mutate(`Family Level Excluded Taxa` = 
             ifelse(`Genus Level Excluded Taxa` == -1, 
                    ifelse(`SuperfamilyTaxaLevel` == TRUE | `SuborderTaxaLevel` == TRUE | `OrderTaxaLevel` == TRUE | 
                             `SubclassTaxaLevel` == TRUE | `ClassTaxaLevel` == TRUE , -1, 0), 0 )) %>%
    # had to get super ifelse nesty here to make this logic work, ugly but works
    group_by(BenSampID, `Final VA Family ID`) %>%
    summarise(`Family Level Count` = sum(Count), 
              #`Genus Level Excluded Taxa` = sum(`Genus Level Excluded Taxa`),
              `Family Level Taxa` = n(),
              `Family Level Excluded Taxa` = sum(`Family Level Excluded Taxa`),
              `Final Family Level Taxa` = `Family Level Taxa` + sum(`Family Level Excluded Taxa`) )
  
  # Join bug traits
  bugTraits <- left_join(exclusionMath,vmast,by=c('Final VA Family ID') )
  
  
  if(SCIchoice == 'VSCI'){SCI <- VSCIcalculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VSCI',
           `SCI Threshold` = 60) %>% 
    rename("SCI Score" ="Fam SCI")}
  if(SCIchoice == 'VCPMI + 63'){SCI <- VCPMI63calculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VCPMI + 63',
           `SCI Threshold` = 42) %>% 
    rename("SCI Score" ="CPMI63+CHOWAN")}
  if(SCIchoice == 'VCPMI - 65'){SCI <- VCPMI65calculation(bugTraits,exclusionMath,vmast) %>%
    mutate(SCI = 'VCPMI - 65',
           `SCI Threshold` = 42) %>% 
    rename("SCI Score" ="CPMI65-CHOWAN")}
  
  SCI <- left_join(SCI, benSamps, by = 'BenSampID')
  
  return(SCI) 
}
```


## ODS Data

If you want to pull data from ODS, you need to run this chunk to establish connections. Otherwise the rest of the script will work fine with local data. Only need to set up one pool connection, choose which DB you want to connect to (production or test).

```{r}

## Connect to ODS production
pool <- dbPool(
  drv = odbc::odbc(),
  Driver = "SQL Server Native Client 11.0", 
  Server= "DEQ-SQLODS-PROD,50000",
  dbname = "ODS",
  trusted_connection = "yes"
)

## Connect to ODS_test
# establish db connection locally
#pool <- dbPool(
#  drv = odbc::odbc(),
#  Driver = "SQL Server",  # note the space between SQL and Server ( how MS named driver)
#  Server= "WSQ04151,50000",
#  dbname = "ODS_test"
#)
```




## Master Taxa list

Pull it directly from the server (freshest copy).

```{r}
masterTaxaGenus <- pool %>% tbl("Edas_Benthic_Master_Taxa_View") %>%
      as_tibble() %>%
      # make columns match expected format
      rename('Phylum' = 'PHYLUM_NAME',
             'Class' = 'CLASS_NAME',
             'Subclass' = 'SUBCLASS_NAME',
             'Order' = 'ORDER_NAME',
             'Suborder' = 'SUBORDER_NAME',
             'Superfamily' = 'SUPERFAMILY_NAME',
             'Family' = 'FAMILY_NAME',
             'Subfamily' = 'SUBFAMILY_NAME',
             'Tribe' = 'TRIBE_NAME',
             'Genus' = 'GENUS_NAME',
             'Species' = 'SPECIES_NAME',
             "Final VA Family ID" =  "WBMT_FINAL_FAMILY_ID",
             "FinalID" = "WBMT_FINAL_ID",
             "TolVal" = "WBMT_TOLERANCE_VALUE",
             "FFG" =   "FEEDING_GROUP", 
             "Habit" = "HABIT", 
             "FamFFG" =  "FAMILY_FEEDING_GROUP",
             "FamTolVal" = "WBMT_FAM_TOLERANCE_VALUE",
             "FamHabit" ="FAMILY_HABIT") %>% 
      dplyr::select(Phylum, Class, Subclass, Order, Suborder, Superfamily, Family, Subfamily, Tribe, 
                    Genus, Species, `Final VA Family ID`, FinalID, TolVal, FFG, 
                    Habit, FamFFG, FamTolVal, FamHabit) # keep EDAS Master Taxa list names
```

Or bring it in from a local file (capnia issue FYI).

```{r}
masterTaxaGenus <- read_csv('data/masterTaxaGenus.csv')

```

## Tolerance Value Stuff

Critical to SCI calculation

```{r}
vmast <- masterTaxaGenus %>%
        # get Family level tolerance value, FFG
        rename('GenusTolVal' = 'TolVal',
               'TolVal' = 'FamTolVal',
               'GenusFFG' = 'FFG',
               'FFG' = 'FamFFG',
               'GenusHabit' = 'Habit',
               'Habit' = 'FamHabit') %>%
        mutate(e=ifelse(Order=="Ephemeroptera", 1, 0),
               p=ifelse(Order=="Plecoptera",1,0),
               t=ifelse(Order=="Trichoptera", 1, 0),
               tmin=ifelse((Order=="Trichoptera" & Family != "Hydropsychidae") | 
                             (Order=="Trichoptera" & is.na(Family)) , 1, 0), 
               ept=ifelse(e+p+t>0,1,0), 
               scraper = ifelse(FFG=="Scraper", 1, 0),
               chiro = ifelse(Family=="Chironomidae",1, 0),
               ptmin = ifelse(p + tmin > 0,1,0),
               `clinger-HS` = ifelse(Habit == 'Clinger' & ! Family %in% c("Hydropsychidae","Simuliidae"), 1, 0)) %>%
        # Then put it in long format so it can be merged to and input taxa list
        select(`Final VA Family ID`,TolVal, e,p,t, ept,ptmin, scraper, chiro,`clinger-HS`) %>% 
        distinct(`Final VA Family ID`, .keep_all = T) %>% # drop multiple rows bc working back to family level data from genus
        filter(!is.na(`Final VA Family ID`)) %>%
        pivot_longer(-`Final VA Family ID`, names_to = 'metric', values_to = 'metric_val') %>%
        #  pivot_longer(-`Final VA Family ID`, names_to = 'metric', values_to = 'metric_val') %>%
        filter(!is.na(metric_val)) 
```


## Sample data 

This shows you how to pull sample data from one station from you pool object. YOu can alter to pull more that one at a time.

First establish station(s).

```{r}
station <- '2-JKS023.61'
```


```{r}
stationBenSamps <- pool %>% tbl("Edas_Benthic_Sample_View") %>%
        filter(STA_ID %in% !! toupper(station)) %>%
        as_tibble() %>%
        # fix names
        rename( "StationID" = "STA_ID",
                "BenSampID"  = "WBS_SAMP_ID",
                "RepNum" = "WBS_REP_NUM",
                "Sample Comments" = "WBS_COMMENT",
                "Entered By" = "WBS_INSERTED_BY", # not in EDAS table but good info
                "Collected By" = "COLLECTOR_NAME",  # not in EDAS table but good info
                "Entered Date" = "WBS_INSERTED_DATE",
                "Gradient" = "WBCM_DESCRIPTION",
                "Taxonomist" = "TAXONOMIST_NAME",  # not in EDAS table but good info
                "Target Count" = "WBS_TARGET_COUNT",
                "Field Team" = "WBS_FIELD_TEAM",
                "Collection Date" = "FDT_DATE_TIME") %>%
        # Add sample season 
        mutate(monthday = as.numeric(paste0(sprintf("%02d",month(`Collection Date`)),
                                            sprintf("%02d",day(`Collection Date`)))),
               Season = case_when(monthday >= 0215 & monthday <= 0615 ~ 'Spring',
                                  monthday >= 0815 & monthday <= 1215 ~ 'Fall',
                                  TRUE ~ as.character("Outside Sample Window"))) %>%
        dplyr::select(StationID, BenSampID, RepNum, `Collection Date`, `Sample Comments`, `Collected By`, `Field Team`, `Entered By`,
                      Taxonomist, `Entered Date`, Gradient, `Target Count`, Season)
```

If you want this from local data, match the below format.

```{r}
stationBenSamps <- read_csv('data/stationInfoBenSampsTESTSITE.csv')
```

## Benthic Data

Pull station's benthic data from server.

```{r}
stationBenthics <- pool %>% tbl("Edas_Benthic_View") %>%
        filter(STA_ID %in% !! toupper(station)) %>%
        as_tibble() %>%
        rename( "StationID" = "STA_ID",
                "BenSampID"  = "WBS_SAMP_ID",
                "RepNum" = "WBS_REP_NUM",
                "FinalID" = "WBMT_FINAL_ID",
                "Individuals" = "WBE_INDIVIDUALS",
                "ID Comments" = "WBE_COMMENT",
                "Entered By" = "WBE_INSERTED_BY", # not in EDAS table but good info
                "Taxonomist" = "TAXONOMIST_NAME",  # not in EDAS table but good info
                "Entered Date" = "WBE_INSERTED_DATE") %>%
        mutate(`Excluded Taxa` = ifelse(WBE_EXCLUDED_TAXA_YN == "Y", -1, 0)) %>%
        dplyr::select(StationID, BenSampID, RepNum, FinalID, Individuals, `Excluded Taxa`, `ID Comments`, Taxonomist, `Entered By`, `Entered Date`)
```


Or filter out benthic data from benthic dataset pulled 11/13/2020.

```{r}
stationBenthics <- read_csv('data/benthics11132020.csv') %>%
  filter(StationID %in% station)
```


## SCI

If you want to limit samples you want to run, do it now, otherwise do it after. The SCI() function let's you choose which method you want to run samples with.

```{r}
# Run data with VSCI
SCIresults <- SCI(stationBenthics, "VSCI", stationBenSamps, masterTaxaGenus, vmast) %>%
        mutate_if(is.numeric, round, digits=2) %>%# rounds all numeric columns to 2 decimal places
        arrange(`Collection Date`)

# Run data with VCPMI + 63
SCIresults <- SCI(stationBenthics, "VCPMI + 63", stationBenSamps, masterTaxaGenus, vmast) %>%
        mutate_if(is.numeric, round, digits=2) %>%# rounds all numeric columns to 2 decimal places
        arrange(`Collection Date`)

# Run data with VCPMI - 65
SCIresults <- SCI(stationBenthics, "VCPMI - 65", stationBenSamps, masterTaxaGenus, vmast) %>%
        mutate_if(is.numeric, round, digits=2) %>%# rounds all numeric columns to 2 decimal places
        arrange(`Collection Date`)
```

      